Staged Abstract Interpreters

### 2019-02-20 paper/10_intro.tex
misc

Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close
connection between compilers and interpreters. The first Futamura projection
specifically shows that specializing an interpreter with respect to the input
program yields an equivalent executable. Partial evaluation
\cite{DBLP:books/daglib/0072559} was the first proposed approach to realize
Futamura projections, which first identifies the binding-time of variables in
the program, i.e., they can be known whether statically or dynamically, and then
evaluates the static part, and finally generates a residual program that solely
relies on the dynamic part. However, given an arbitrary program, precisely
analyzing its binding-time is hard in general. As an alternative and pragmatic
approach to specialization and partial evaluation, multi-stage programming (MSP
for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage
annotations (i.e., binding-time annotations) to be explicit from the programmers.
These staging annotations identify which part of the inputs should be
specialized. As a classical example, we consider the power function to introduce
the idea of MSP:

\begin{lstlisting}
def power(b: Rep[Int], e: Int): Int = 
  if (e == 0) 1 else e * power(b, e-1)
\end{lstlisting}

@@

Roughly at the same time when Futamura projections was visioned in 1970s,
\citeauthor{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
lattice-based semantic approach to construct sound static analysis, by
approximation of fixpoints \cite{DBLP:conf/popl/CousotC77}. However,
constructing sound abstract interpreters was considered abstruse and complicated
for a long time.
Recent progress such as Abstracting Abstract Machines (AAM)
uncovers a systematic method to derive sound abstract interpreters from their
concrete counterparts, where the soundness can be easily established by
examining the transformation of semantics.
The AAM approach also has been been applied to different variants of
definitional interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
\todo{CEK example?}

>>>>

Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close
connection between compilers and interpreters. The first Futamura projection
specifically shows that specializing an interpreter with respect to the input
program yields an equivalent executable. Partial evaluation
\cite{DBLP:books/daglib/0072559} was the first proposed approach to realize
Futamura projections, which first identifies the binding-time of variables in
the program, i.e., they can be known whether statically or dynamically, and then
evaluates the static part, and finally generates a residual program that solely
relies on the dynamic part. However, given an arbitrary program, precisely
analyzing its binding-time is hard in general. As an alternative and pragmatic
approach to specialization and partial evaluation, multi-stage programming (MSP
for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage
annotations (i.e., binding-time annotations) to be explicit from the programmers.
These staging annotations identify which part of the inputs should be
specialized. As a classical example, we use the power function to introduce
the idea of MSP:

\begin{lstlisting}
def power(b: Rep[Int], e: Int): Rep[Int] = 
  if (e == 0) 1 else e * power(b, e-1)
\end{lstlisting}

@@

Roughly at the same time when Futamura projections was visioned in 1970s,
\citeauthor{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
lattice-based semantic approach to construct sound static analysis, by
approximation of fixpoints \cite{DBLP:conf/popl/CousotC77}. However,
constructing sound abstract interpreters was considered abstruse and complicated
for a long time.
Recent progress such as Abstracting Abstract Machines (AAM)
uncovers a systematic method to derive sound abstract interpreters from their
concrete counterparts, where the soundness can be easily established by
examining the transformation of semantics.
The AAM approach also has been been applied to different variants of
definitional interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
% \todo{CEK example?}
